--- 
title: "PD Stream patient progress report"
author: "HWU"
date: "`r format(Sys.time(), '%d %B, %Y @ %H:%M')`"
output: pdf_document
papersize: a4
header-includes:    
    - \usepackage{pdflscape}
---

```{r setup, include=FALSE}
### Variables:

linePath = TRUE
lineSmooth = FALSE
showStdError = FALSE
spanValue = 0.3

# span controls line smoothness - if smooth is chosen. Usually you don't want it

### Libraries:

library(limer)
library(properties)
library(cowplot)
library(knitr)
library(lubridate)
library(tidyverse)
library(scales)
# if extra nice PDF output wanted use tikzDevice
# library(tikzDevice)
# then in plot rendering chunks add option, dev='tikz'
# then must \\ double-slash escape % and & in plot titles
# rendering will take slightly longer


### Begin:

# read properties - aka parameters / variables
props <- read.properties("script.properties")

# knitr::opts_knit$set(root.dir = './out/')

if (props$onePerson=="YES") {
  onePerson <- TRUE
  onePersonNumber <- as.integer(props$onePersonNumber)
  if (is.na(onePersonNumber)) {
    stop('You want graphs for one patient, but patient number is wrong or missing \n
         Please check the .properties file to see if variable onePersonNumber is set')
    }  
  } else {
    onePerson <- FALSE
  }


startDT <- ymd_hm(paste(props$dateStart, props$timeStart))
endDT <- ymd_hm(paste(props$dateEnd, props$timeEnd))



if (is.na(startDT)) {
    stop('Start date or Time not in the right format in the properties file')
  }  else if (is.na(endDT)) {
    stop('Ending date or Time not in the right format in the properties file')  
  }


### Reading in data
# GET data

options(lime_api = props$limeAPI)
options(lime_username = props$limeUser)
options(lime_password = props$limePassword)

### Reading in data

skey <- get_session_key()
if(skey == '') { stop("No connection to database. Is network running OK?")}

initData <- get_responses(props$limeSurveyNumber)

if (props$resetDemo == "YES") {
  source("./resetDemoData.R")
  print("DEMO data updated, reacquiring initial dataset")
  initData <- get_responses(props$limeSurveyNumber)
}

# clear session with limesurvey, do not print output of this (via invisible)
invisible(release_session_key())



# remove trailing dot from column names and set date datatype on submitdate
names(initData) <- gsub("\\.$", "", names(initData))
initData <- initData %>% mutate(submitdate = ymd_hms(submitdate))
initData <- initData %>% mutate(date = as_date(submitdate))
initData <- initData %>% arrange(submitdate)
initData <- initData %>% mutate(mdate = 
             floor_date(date, unit = "week", week_start = getOption("lubridate.week.start", 1)))


# proceed to rearrange data as needed
df.lines.feeling <- initData %>%
  select(date,mdate,id=Q003ID,feeling=Q001SL.SQ001) %>% 
  filter(date >= startDT) %>% 
  filter(date <= endDT)
if(onePerson) { df.lines.feeling <- df.lines.feeling %>% 
  filter(., id==onePersonNumber)
  }

df.avg.feeling <- df.lines.feeling %>%
  dplyr::group_by(mdate) %>%
  summarise(feeling=mean(feeling))

# PLOT for Feeling lines
p.lines.individual <- ggplot(df.lines.feeling,aes(x=mdate,y=feeling, colour=as.factor(id))) +
  {if (lineSmooth) stat_smooth(geom="line",aes(x=mdate,y = feeling, color = as.factor(id)), span = spanValue, se = showStdError, alpha = 0.5)} +
  {if (linePath) geom_path(aes(x = mdate, y = feeling, color = as.factor(id)))} +
  {if (!onePerson) geom_line(data=df.avg.feeling,aes(x=mdate,y=feeling,colour="AVG for Unit"),alpha = 0.8, size = 3)} +
  labs(title = paste("Feeling over time on 100% scale from",
                     format(as_datetime(min(initData$submitdate)),"%b %Y"),
                     "until",
                     format(as_datetime(max(initData$submitdate)),"%b %Y")
                     ), x = NULL, y = "feeling", color="Patients") +
  scale_y_continuous(breaks=pretty_breaks(), limits=c(0,100)) +
  theme_minimal()


df.lines.emergencies <- initData %>% 
  select(date,mdate,id=Q003ID,ae=Q002MC.SQ001,police=Q002MC.SQ002,apts=Q002MC.SQ003) %>% 
  filter(date >= startDT) %>% 
  filter(date <= endDT)
if(onePerson) {df.lines.emergencies <- df.lines.emergencies %>% 
  filter(id==onePersonNumber)
  }

df.avg.ae <- df.lines.emergencies %>%
  select(date,mdate,ae) %>%
  group_by(mdate) %>%
  summarise(ae=mean(ae))

# get top scale limit for Y axis for emergencies
maxEmergencyNumber <- df.lines.emergencies %>% select(ae,police,apts) %>% max(.,na.rm = TRUE)
if (maxEmergencyNumber <=10) {
  maxEmergencyNumber <- 10
}

# PLOT for A&E attendance lines
p.lines.ae <- ggplot(df.lines.emergencies,aes(x=mdate,y=ae,colour=as.factor(id))) +
  {if (lineSmooth) stat_smooth(geom="line",aes(x=mdate,y = ae, color = as.factor(id)), span = spanValue, se = showStdError, alpha = 0.5)} +
  {if (linePath) geom_path(aes(x = mdate, y = ae, color = as.factor(id)),alpha=0.6)} +
  {if (!onePerson) geom_line(data=df.avg.ae,aes(x=mdate,y=ae,colour="AVG for Unit"),alpha=0.8,size=3)} +
  labs(title = "A&E attendance / admissions", x=NULL, y="A\\&E or admissions") +
  scale_y_continuous(breaks=seq(0,maxEmergencyNumber,2), limits=c(0,maxEmergencyNumber)) +
  theme_minimal() +
  theme(legend.position = "none")


df.avg.police <- df.lines.emergencies %>%
  select(date,mdate,police) %>%
  group_by(mdate) %>%
  summarise(police=mean(police))

# PLOT for Police contact lines
p.lines.police <- ggplot(df.lines.emergencies,aes(x=mdate,y=police,colour=as.factor(id))) +
  {if (lineSmooth) stat_smooth(geom="line",aes(x=mdate,y = police, color = as.factor(id)), span = spanValue, se = showStdError, alpha = 0.6)} +
  {if (linePath) geom_path(aes(x = mdate, y = police, color = as.factor(id)),alpha=0.6)} +
  {if (!onePerson) geom_line(data=df.avg.police,aes(x=mdate,y=police,colour="AVG for Unit"),alpha=0.8,size=3)} +
  labs(title = "Police contacts", x=NULL, y="Police contact") +
  scale_y_continuous(breaks=seq(0,maxEmergencyNumber,2), limits=c(0,maxEmergencyNumber)) +
  theme_minimal() +
  theme(legend.position = "none")


df.avg.apts <- df.lines.emergencies %>%
  select(date,mdate,apts) %>%
  group_by(mdate) %>%
  summarise(apts=mean(apts))

# PLOT for Missed Appointments lines (this has legend!)
p.lines.apts <- ggplot(df.lines.emergencies,aes(x=mdate,y=apts,colour=as.factor(id))) +
  {if (lineSmooth) stat_smooth(geom="line",aes(x=mdate,y = apts, color = as.factor(id)), span = spanValue, se = showStdError, alpha = 0.6)} +
  {if (linePath) geom_path(aes(x = mdate, y = apts, color = as.factor(id)),alpha=0.6)} +
  {if (!onePerson) geom_line(data=df.avg.apts,aes(x=mdate,y=apts,colour="AVG for Unit"),alpha=0.8,size=3)} +
  labs(title = "Crisis contacts", x=NULL, y="Number of CRHT", color = "Patients") +
  scale_y_continuous(breaks=seq(0,maxEmergencyNumber,2), limits=c(0,maxEmergencyNumber)) +
  theme_minimal() 


# PLOT 3in1 - put together three service (mis)use indicators
p.3in1 <- plot_grid(p.lines.ae,p.lines.police,p.lines.apts, labels = "AUTO", nrow = 1, ncol = 3, align = "vh", axis = "tb", rel_widths = c(1,1,1.3))
```

# HWU Weekly Coping estimate 

We are asking our patients at the beginning of each individual session (typically weekly or fortnightly) to tell us how they have felt overall for the past week on scale from 0 to 100 %. We also ask how many times they have been in contact with emergency or crisis services. We think that patients' recollection/impression of the last 7 days is fairly accurate. 

\vspace{15mm}

```{r PlotFeeling, echo=F, warning=F, error=F, message=F}
p.lines.individual

```



\newgeometry{left=0.5in,right=0.5in,top=0.5in,bottom=0.5in,nohead}
\begin{landscape}
```{r Plot3in1, echo=F, warning=F,error=F, message=F, fig.align='center', fig.width = 26/2.54, fig.height = 18/2.54}

# page margins needed to be decreased to avoid blank page before the 3in1 plot

p.3in1

```
\end{landscape}
